```{r, echo=FALSE, include=FALSE}
## Get VERSION and create file names
ver <- sub(x=grep(x=readLines("DESCRIPTION"), pattern="^Version: ",
                  value=TRUE),
           pattern="^Version: ", replacement="")
pkg.name <- "mcglm_"
pkg.source <- paste0(pkg.name, ver, ".tar.gz")
pkg.win <- paste0(pkg.name, ver, ".zip")
```

# mcglm `r ver` 

[![Build Status](https://travis-ci.org/wbonat/mcglm.svg?branch=master)](https://travis-ci.org/wbonat/mcglm)
Build status for the stable version (`master` branch)

[![Build Status](https://travis-ci.org/wbonat/mcglm.svg?branch=devel)](https://travis-ci.org/wbonat/mcglm)
Build status for the development version (`devel` branch)

The `mcglm` package fits multivariate covariance generalized linear models (McGLMs).

## Introduction

McGLM is a general framework for non-normal multivariate data analysis, designed 
to handle multivariate response variables, along with a wide range of temporal 
and spatial correlation structures defined in terms of a covariance link function 
combined with a matrix linear predictor involving known matrices. 
The models take non-normality into account in the conventional way by means of a 
variance function, and the mean structure is modelled by means of a link function 
and a linear predictor. The models are fitted using an efficient Newton scoring 
algorithm based on quasi-likelihood and Pearson estimating functions, using only 
second-moment assumptions. This provides a unified approach to a wide variety of 
different types of response variables and covariance structures, including 
multivariate extensions of repeated measures, time series, longitudinal, spatial 
and spatio-temporal structures. The package offers a user-friendly interface for 
fitting McGLMs similar to the `glm()` `R` function. 
See [Bonat (2018)](https://www.jstatsoft.org/article/view/v084i04), for more 
information and examples.

## Download and install

### Linux/Mac

Use the `devtools` package (available from
[CRAN](http://cran-r.c3sl.ufpr.br/web/packages/devtools/index.html)) to
install automatically from this GitHub repository:

```{r, eval=FALSE}
library(devtools)
install_github("wbonat/mcglm")
```

Alternatively, download the package tarball: [`r pkg.source`][]
and run from a UNIX terminal (make sure you are on the container file
directory):

```{r, echo=FALSE, comment=NA}
cmd <- paste("R CMD INSTALL -l /path/to/your/R/library", pkg.source)
cat(cmd, sep = "\n")
```

Or, inside an `R` session:

```{r, echo=FALSE, comment=NA}
inst <- paste0("install.packages(", "\"", pkg.source, "\"", ", repos = NULL,\n",
               "                 lib.loc = \"/path/to/your/R/library\",\n",
               "                 dependencies = TRUE)")
cat(inst, sep = "\n")
```

Note that `-l /path/to/your/R/library` in the former and `lib.loc =
"/path/to/your/R/library"` in the latter are optional. Only use it if
you want to install in a personal library, other than the standard R
library.

### Windows

Download Windows binary version: [`r pkg.win`][] (**do not unzip
it under Windows**), put the file in your working directory, and from
inside `R`:

```{r, echo=FALSE, comment=NA}
instw <- paste0("install.packages(", "\"", pkg.win, "\"", ", repos = NULL,\n",
                "                 dependencies = TRUE)")
cat(instw, sep = "\n")
```

### Development version

By default, if you use `devtools::install_github()`, or download any of the
package tarball or Windows binary version, it will install the stable
version of the package (from the `master` branch of this repository).

If you want to install the development version, you can use

```r
library(devtools)
install_github("wbonat/mcglm", ref = "devel")
```

Note that the development version can contain bugs and other unknown
features, so use it at your own risk!

## Authors

- [Wagner Hugo Bonat][] (author and main developer)
- [Walmes Marques Zeviani][] (contributor)
- [Fernando de Pol Mayer][] (contributor)

## Documentation

The reference manual in PDF can be found here: [mcglm-manual.pdf][]

## Contributing

This R package is develop using [`roxygen2`][] for documentation and
[`devtools`] to check and build. Also, we adopt the [Gitflow worflow][]
in this repository. Please, see the
[instructions for contributing](./CONTRIBUTING.md) to collaborate.

## License

This package is released under the
[GNU General Public License (GPL) v3.0][].

See [LICENSE](./LICENSE)

<!-- links -->

```{r, echo=FALSE, include=FALSE}
pkg.source.link <- paste0(
    "https://github.com/wbonat/mcglm/raw/master/downloads/",
    pkg.source)
pkg.win.link <- paste0(
    "https://github.com/wbonat/mcglm/raw/master/downloads/",
    pkg.win)
```

[GNU General Public License (GPL) v3.0]: http://www.gnu.org/licenses/gpl-3.0.html
[`roxygen2`]: https://github.com/klutometis/roxygen
[`devtools`]: https://github.com/hadley/devtools
[`r pkg.source`]: `r pkg.source.link`
[`r pkg.win`]: `r pkg.win.link`
[mcglm-manual.pdf]: http://www.leg.ufpr.br/~wagner/docmcglm/mcglm_0.6.0.pdf
[Gitflow worflow]: http://nvie.com/posts/a-successful-git-branching-model/
[Wagner Hugo Bonat]: http://www.leg.ufpr.br/~wagner
[Walmes Marques Zeviani]: http://www.leg.ufpr.br/~walmes
[Fernando de Pol Mayer]: http://www.leg.ufpr.br/~fernandomayer
